Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 2.dev

### 2024-09-19 main.tex
updated the captions and added definition of perturbation order.

\subsection{Measurements}
%
We measure the time histories of the Bump'Em delivered perturbation forces and
the bicycle's steer angle, roll angle, roll angular rate, and forward speed.
Figure~\ref{fig:perturbation_0} shows an example perturbation force measurement
compared to our Bump'Em controller command.
%
\begin{figure}
  \centering
  \includegraphics[width=140mm]{figures/torque_angle_perturbation_10.png}
  \caption{Perturbation torque alongside the resulting motion and steer motor
    torque based on a 110~\si{\newton} counterclockwise applied force at a a
    6~\si{\kph} travel speed.}
  \label{fig:perturbation_0}
\end{figure}

@@

At the initiation of each perturbation we log the instantaneous steer and roll
angles to characterize the configuration of the bicycle when perturbed. The
gain setting on the balance assist controller indicates if the assistance is
off \(g=0\) or on at two different levels: low \(g=8\) or high \(g=10\). A
recovery from the perturbation is successful if the person neither places their
foot down onto the treadmill surface nor their wheel of the bicycle exits the
width of the treadmill belt. We record this as a binary variable \(f\) for
"fall". All measured variables are reported in Table~\ref{tab:raw}.
%
\begin{table}
  \centering
  \caption{Raw measurements taken during each trial. ``Fall outcome'' and
    ``Perturbation order number'' are recorded per perturbation instance. The
    remaining measures are time varying during each trial.}
  \begin{tabular}{llll}
    \toprule
    Measure & Variable & Units & Sensor \\
    \midrule
    Balance Assist Gain & \(g\) & \si{\newton\second\squared} & NA \\
    Bicycle Speed & \(v\) & \si{\meter\per\second} & wheel encoder \\
    Fall outcome & \(f\) & boolean & NA \\
    Force \(l\)eft/\(r\)ight,\(f\)ront/\(r\)ear &
    \(F_{lf},F_{rf},F_{lr},F_{rr}\) & \si{\newton} & inline load cell\\
    Perturbation Order Number & \(j\) & integer & NA\\
    Roll Angle & \(\phi\) & \si{\degree} & BN0086 Kalman estimate \\
    Roll Angular Rate & \(\dot{\phi}\) & \si{\degree\per\second} & BN0086 rate gyroscope \\
    Steer Angle & \(\delta\) & \si{\degree} & steer tube encoder \\
    \bottomrule
  \end{tabular}
  \label{tab:raw}
\end{table}

@@

We evaluate this hypothesis using a multivariate logistic regression model that
takes the general form
%
\begin{align}
  f_{ij} | p_{ij} \sim \textrm{Bern}(p_{ij}) \textrm{.}
\end{align}
%
Fall outcome \(f_{ij}\) is the binary outcome of perturbation \(j\) on
participant \(i\) which follows a Bernoulli distribution given the probability
\(p_{ij}\) that a fall occurs. The log-odds of the probability is then a linear
function of our independent variables with \(\beta\) being the intercept and
\(\alpha_k\) the linear coefficients to the \(K\) independent variables
\(x_{ij}^k\), i.e. all variables in Table~\ref{tab:stat-model-variables} except
\(f\).
%
\begin{align}
  \log \left(\frac{p_{ij}}{1-p_{ij}} \right) =
  \beta + \sum_{k=0}^{K} \alpha_k x_{ij}^{k}
  \label{eq:log-regress}
\end{align}

@@

\section*{Acknowledgements}
%
This study follows and draws from experimental and analysis methods originally
developed in unpublished research by Marco Reijne. The authors acknowledge
Felix Dauer, David Gabriel, Sierd Heida, Oliver Maier, Maarten Pelgrim, Marco
Reijne, and Arend Schwab for contributions to the development of the balance
assist bicycle. We thank all of the assistants that helped run the experiments
and the subjects for their time an enthusiasm.

>>>>

\subsection{Measurements}
%
We measure the time histories of the Bump'Em delivered perturbation forces and
the bicycle's steer angle, roll angle, roll angular rate, and forward speed.
Figure~\ref{fig:perturbation_0} shows an example perturbation force measurement
compared to our Bump'Em controller command.
%
\begin{figure}
  \centering
  \includegraphics[width=140mm]{figures/torque_angle_perturbation_10.png}
  \caption{Externally applied perturbation torque alongside the resulting
    measured motion and steer motor induced torque based on a 110~\si{\newton}
    counterclockwise applied force at a 6~\si{\kph} travel speed. The circles
    on the roll and steer angle plot show the angles \(\phi_0\) and
    \(\delta_0\) at the perturbation start.}
  \label{fig:perturbation_0}
\end{figure}

@@

We record the order number of each perturbation, i.e. first, second, third,
\ldots, to measure how many perturbations a rider is subjected to before the
current pertubration. At the initiation of each perturbation we log the
instantaneous steer and roll angles to characterize the configuration of the
bicycle when perturbed. The gain setting on the balance assist controller
indicates if the assistance is off \(g=0\) or on at two different levels: low
\(g=8\) or high \(g=10\). A recovery from the perturbation is successful if the
person neither places their foot down onto the treadmill surface nor their
wheel of the bicycle exits the width of the treadmill belt. We record this as a
binary variable \(f\) for "fall". All measured variables are reported in
Table~\ref{tab:raw}.
%
\begin{table}
  \centering
  \caption{Raw measurements taken during each trial. ``Fall outcome'' and
    ``Perturbation order number'' are recorded per perturbation instance. The
    remaining measures are time varying during each trial.}
  \begin{tabular}{llll}
    \toprule
    Measure & Variable & Units & Sensor \\
    \midrule
    Balance Assist Gain & \(g\) & \si{\newton\second\squared} & NA \\
    Bicycle Speed & \(v\) & \si{\meter\per\second} & wheel encoder \\
    Fall outcome & \(f\) & boolean & NA \\
    Force \(l\)eft/\(r\)ight,\(f\)ront/\(r\)ear &
    \(F_{lf},F_{rf},F_{lr},F_{rr}\) & \si{\newton} & inline load cell\\
    Perturbation Order Number & \(j\) & integer & NA\\
    Roll Angle & \(\phi\) & \si{\degree} & BN0086 Kalman estimate \\
    Roll Angular Rate & \(\dot{\phi}\) & \si{\degree\per\second} & BN0086 rate gyroscope \\
    Steer Angle & \(\delta\) & \si{\degree} & steer tube encoder \\
    \bottomrule
  \end{tabular}
  \label{tab:raw}
\end{table}

@@

Fall outcome \(f_{ij}\) is the binary outcome of perturbation \(j\) on
participant \(i\) which follows a Bernoulli distribution given the probability
\(p_{ij}\) that a fall occurs:
%
\begin{align}
  f_{ij} | p_{ij} \sim \textrm{Bernoulli}(p_{ij}) \textrm{.}
\end{align}
%
We evaluate our hypothesis using a multivariate logistic regression model that
represents the probability of this binary outcome. The log-odds of the
probability is then a linear function of our independent variables with
\(\beta\) being the intercept and \(\alpha_k\) the linear coefficients to the
\(K\) independent variables \(x_{ij}^k\), i.e. all variables in
Table~\ref{tab:stat-model-variables} except \(f\).
%
\begin{align}
  \log \left(\frac{p_{ij}}{1-p_{ij}} \right) =
  \beta + \sum_{k=0}^{K} \alpha_k x_{ij}^{k}
  \label{eq:log-regress}
\end{align}

@@

\section*{Acknowledgements}
%
This study follows and draws from experimental and analysis methods originally
developed in unpublished research by Marco Reijne. The authors acknowledge
Felix Dauer, David Gabriel, Sierd Heida, Oliver Maier, Maarten Pelgrim, Marco
Reijne, and Arend Schwab for contributions to the development of the balance
assist bicycle. We thank all of the assistants that helped run the experiments
and the subjects for their time an enthusiasm. We thank Peter Stahlecker for
valuable feedback in early revisions.
